---
title: "Homework 4"
author: "Peyton Hall"
date: "02/14/2025"
output:
  pdf_document:
    latex_engine: lualatex
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load Necessary Libraries
```{r Load Necessary Libraries}
library(readxl)
library(tidyverse)
library(ggplot2)
```


1. An obstetrician wanted to determine the impact that two experimental diets 
(A and B) and age of moms had on the birth weights of the infants. She randomly 
selected 18 pregnant mothers in the first trimester of whom 6 were 20 to 29 
years old, 6 were 30 to 39 years old, and 6 were 40 or older. For each age 
group, she randomly assigned the mothers to one of the two diets. After delivery 
she measured the birth weights (in grams) of the babies and obtained the 
following data. Assume that the data follow the normal distribution.
Age,             Diet
               A      B
20 - 29,     4473,   3667
             3878,   3139
             3936,   3356
30 - 39,     3886,   2762
             4147,   3551
             3693,   3272
40 or older, 3878,   2781
             4002,   3138
             3382,   3435
a) What is the null and alternative hypothesis, in symbols, for testing the main 
effect of age?
$H_0: μ_\text{20−29} = μ_\text{30−39} = μ_\text{40+}$ vs $H_a: \text{At least one μ differs}$
b) What is the test statistic and p-value for testing the main effect of age?
c) What conclusion in the context can you draw based on the p-value in b)? (is 
there any significant mean weight difference between the three age groups?)
d) What is the null and alternative hypothesis, in symbols, for testing the main 
effect of diet?
$H_0: μ_\text{DietA} = μ_\text{DietB}$ vs $H_a: μ_\text{DietA} ≠ μ_\text{DietB}$
e) What is the test statistic and p-value for testing the main effect of diet?
f) What conclusion in the context can you draw based on the p-value in e)? (is 
there any significant mean weight difference between the A and B diet groups?)
g) Is there a significant interaction between diet and age? List the p-value to 
explain.

Question 1 Code
```{r Question 1 Code}
# b) find test statistic and p-value
age_group <- rep(c("20 - 29", "30 - 39", "40 or older"), each = 3)
diet_a <- c(4473, 3878, 3936, 3886, 4147, 3693, 3878, 4002, 3382)
diet_b <- c(3667, 3139, 3356, 2762, 3551, 3272, 2781, 3138, 3435)
diet_data <- data.frame(Age = age_group, Diet_A = diet_a, Diet_B = diet_b)

# reshape data to long format
data_long <- pivot_longer(diet_data, cols = c(Diet_A, Diet_B), 
                          names_to = "Diet", values_to = "Birth_Weight")
data_long$Diet <- factor(data_long$Diet)
data_long$Age <- factor(data_long$Age)

anova_result1 <- aov(Birth_Weight~Age * Diet, data = data_long)
summary(anova_result1)

```
f = 1.416; p-value = 0.280557
Fail to reject H0; there is no evidence to support the claim that there is 
significant mean weight difference between the three age groups.
f = 21.005; p-value = 0.000629
Reject H0; there is significant evidence to support the claim that significant 
mean weight difference between the A and B diet groups exist.
answer part g
f = 0.028; p-value = 0.972449
Fail to reject H0; there is no significant evidence to support the claim that 
there is a significant interaction between diet and age.


2. A researcher took 64 water samples from a stream running through a forest. 
The samples were collected over the course of a year. Each sample was analyzed 
for concentration of dissolved organic carbon (mg/L) in it. Each sample was 
categorized according to the type of water collected: groundwater from organic 
soil (Organic) or groundwater from mineral soil (Mineral), and categorized 
according to the location of the sample: the surface of the stream or the bottom 
of the stream.
The researchers wanted to determine if the mean concentration of dissolved 
organic carbon was the same between the two types of water collected, and 
between the two locations. The researchers want to know which type of water has 
the highest dissolved organic carbon. The data is on D2L. Answer the following 
questions:
a) What are the null and alternative of testing the mean difference between the 
two types of water?
$H_0: μ_\text{Organic} = μ_\text{Mineral}$ vs $H_a: μ_\text{Organic} ≠ μ_\text{Mineral}$
b) What is the test statistic and p-value for a)?
c) What conclusion in the context from the p-value in a)?
d) What are the null and alternative of testing the mean difference between the 
two locations?
$H_0: μ_\text{Surface} = μ_\text{Bottom}$ vs $H_a: μ_\text{Surface} ≠ μ_\text{Bottom}$
e) What is the test statistic and p-value for d)?
f) What conclusion in the context from the p-value in d)?
g) What is the test statistic and p-value for testing the interaction? What can 
you conclude from that p-value?
h) Generate an interaction plot. Based on the plot, which type of water at which 
location reached the highest dissolved organic carbon?

Question 2 Code
```{r Question 2 Code}
SoilTesting <- read_excel("~/Desktop/STAT 301/Week 4/SoilTesting.xls")
SoilTesting$Carbon<-as.numeric(SoilTesting$Carbon)
# b)
organic <- SoilTesting[SoilTesting$Source == "organic", ]
mineral <- SoilTesting[SoilTesting$Source == "mineral", ]
t.test(organic$Carbon, mineral$Carbon, alternative = "two.sided", var.equal = TRUE)
# e)
surface <- SoilTesting %>% filter(Location == "surface")
bottom <- SoilTesting %>% filter(Location == "bottom")
t.test(surface$Carbon, bottom$Carbon, alternative = "two.sided", var.equal = TRUE)

# g)
interaction_result <- aov(Carbon~Source * Location, data = SoilTesting)
summary(interaction_result)

# h)
ggplot(SoilTesting, aes(x = Source, y = Carbon, color = Location, group = Location)) +
  geom_line(aes(linetype = Location)) +
  geom_point() +
  labs(title = "Interaction Plot", x = "Source of Water", y = "Dissolved Organic Carbon") +
  theme_minimal()

```
b and c) f = 2.7642; p-value = 0.007504
Reject h0; there is significant evidence to support the claim that the mean of
the surface does not equal the mean of the bottom
e and f) 
t = -0.41519; p-value = 0.6794
Fail to reject H0; there is significant evidence to support the claim that the
mean of the surface equals the mean of the bottom.
Question 1 Part g):
f = 3.230; p = 0.07734
There is no significant evidence to support the claim that the interaction 
between the source of the water (organic vs. mineral) and the location (surface 
vs. bottom) significantly affects the dissolved organic carbon levels.


3. Researchers have sought to examine the effect of various types of music on 
agitation levels in patients who are in the early and middle stages of 
Alzheimer’s disease. Patients were selected to participate in the study based on 
their stage of Alzheimer’s disease. Three forms of music were tested: Easy 
listening, Mozart, and Piano Interludes. While listening to music, agitation 
levels were recorded for the patients with a high score indicating a higher 
level of agitation. Scores are recorded below. Use the data to answer the 
following questions:
              Piano Interlude, Mozart, Easy Listening
Early State
Alzheimer’s,  21,              9,      29
              24,              12,     26
              22,              10,     30
              18,              5,      24
              20,              9,      26
Middle State
Alzheimer’s,  22,              14,     15
              20,              18,     18
              25,              11,     20
              18,              9,      13
              20,              13,     19
a) What is the test statistic and p-value for testing the mean score difference 
between the early state Alzheimers and the middle state Alzheimers?
b) What do you conclude in the context from the p-value in a)? (Is there any 
significant mean score difference between the two states?)
c) What is the test statistic and p-value for testing the mean score difference 
between the three music types?
d) What can you conclude from the p-value in c)? (Is there any significant mean 
score difference between the three types of music?)
e) Is there a significant interaction between music type and the state of 
Alzheimer? List the p-value.
f) If there is a significant interaction between the music type and the state of 
Alzheimer, separate the data based on the Alzheimer state and answer the 
following:
    a. For the Early State Alzheimer’s, what is the test statistic and p-value 
    to compare the mean scores from the three music types? What can you conclude 
    from the p-value?
    b. For the Middle State Alzheimer’s, what is the test statistic and p-value 
    to compare the mean scores from the three music types? What can you conclude 
    from that p-value?

Question 3 Code
```{r Question 3 Code}
piano_interlude_early <- c(21, 24, 22, 18, 20)
mozart_early <- c(9, 12, 10, 5, 9)
easy_listening_early <- c(29, 26, 30, 24, 26)

piano_interlude_middle <- c(22, 20, 25, 18, 20)
mozart_middle <- c(14, 18, 11, 9, 13)
easy_listening_middle <- c(15, 18, 20, 13, 19)

early_state <- data.frame(
  Piano_Interlude = piano_interlude_early,
  Mozart = mozart_early,
  Easy_Listening = easy_listening_early
)

middle_state <- data.frame(
  Piano_Interlude = piano_interlude_middle,
  Mozart = mozart_middle,
  Easy_Listening = easy_listening_middle
)

# a)
t.test(piano_interlude_early, piano_interlude_middle, var.equal = FALSE)
t.test(mozart_early, mozart_middle, var.equal = FALSE)
t.test(easy_listening_early, easy_listening_middle, var.equal = FALSE)

# b)
# combine the existing data into one dataset
combined_scores <- c(piano_interlude_early, mozart_early, easy_listening_early, 
                     piano_interlude_middle, mozart_middle, easy_listening_middle)
music_type <- factor(rep(c("Piano_Interlude", "Mozart", "Easy_Listening"), each = 10))
stage <- rep(c("Early", "Middle"), each = 15)
# create a data frame for the ANOVA
anova_data <- data.frame(Score = combined_scores, Music_Type = music_type, Stage = stage)

# c) perform ANOVA
anova_results <- aov(Score~Music_Type, data = anova_data)
summary(anova_results)

# e) 
anova_results_interaction <- aov(Score~Music_Type * Stage, data = anova_data)
summary(anova_results_interaction)



```
a) and b):
Piano Interlude:
t = 0.0; p-value = 1.0
Fail to reject H0; there is no evidence to support the claim that there is a 
significant mean agitation score difference between the early and middle state 
Alzheimer's patients when listening to Piano Interlude.
Mozart:
t = -2.1082; p-value = 0.07072
Fail to reject H0; there is no significant evidence to support the claim that 
there is a significant mean agitation score difference between the early and 
middle state Alzheimer's patients when listening to Mozart.
Easy Listening:
t = 5.8722; p-value = 0.0004172
Reject H0; there is significant evidence to support the claim that there is a 
significant mean agitation score difference between the early and middle state 
Alzheimer's patients when listening to Easy Listening music.
c) and d):
f = 10.91; p-value = 0.000336
Reject H0; there is significant evidence to support the claim that significant 
mean score differences between the three music types exist.
e):
Music_Type p-value = 0.000189
Stage: p-value = 0.054681
The interaction term between music type and Alzheimer's stage is not significant.


4. Use the “job and gender” data on d2l. Test the main effects of gender, main 
effects of job category and the interaction between job and gender. Note that 
this is an unbalanced design.
a) Perform a two-way ANOVA to test the main effect of gender
    i) What is the test statistic and p-value for testing the main effect of 
    gender?
    ii) What do you conclude? Is there significant mean salary difference 
    between male and female?
b) Test the main effect of job category
    i) What is the test statistic and p-value for testing the main effect of job 
    category?
    ii) What do you conclude? Is there significant mean salary difference 
    between the janitors, clerks, and managers?
c) If there is a significant interaction, separate the data by job category and 
answer the following:
    a. For the janitors, what is the test statistic and p-value for comparing 
    the mean salary difference between males and females? Is there any 
    significant difference between the average salary from male janitors and 
    female janitors?
    b. For the clerks, what is the test statistic and p-value for comparing the 
    mean salary difference between males and females? Is there any significant 
    difference between the average salary from male clerks and female clerks?
    c. For the managers, what is the test statistic and p-value for comparing 
    the mean salary difference between males and females? Is there any 
    significant difference between the average salary from male managers and 
    female managers?

Question 4 Code
```{r Question 4 Code}
job_and_gender <- read_excel("~/Desktop/STAT 301/Week 4/job and gender.xlsx")

# a)
anova_results <- aov(Salary~Jobcat * Gender, data = job_and_gender)
summary(anova_results)

# c)
janitors <- subset(job_and_gender, Jobcat == "janitor")
clerks <- subset(job_and_gender, Jobcat == "clerk")
managers <- subset(job_and_gender, Jobcat == "manager")

# Perform t-tests comparing male vs. female salaries
t_test_janitors <- t.test(Salary~Gender, data = janitors, var.equal = FALSE)
t_test_clerks <- t.test(Salary~Gender, data = clerks, var.equal = FALSE)
t_test_managers <- t.test(Salary~Gender, data = managers, var.equal = FALSE)
t_test_janitors
t_test_clerks
t_test_managers


```
a)
f = 89.23; p-value = 1.62e-11
There is a significant mean salary difference between male and female.
b)
f = 921.51; p-value < 2e-16
There is a significant mean salary difference between janitors, clerks, and 
managers.
c)
a.
t = -1.4309; p-value = 0.1812
There is no significant mean salary difference between male janitors and female 
janitors.
b.
t = -7.2276; p-value = 3.471e-05
There is a significant mean salary difference between male clerks and female 
clerks.
c.
t = -7.4818; p-value = 2.581e-06
There is a significant mean salary difference between male managers and female 
managers.




