---
title: "Worksheet 4"
author: "Peyton Hall"
date: "02/07/2025"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Libraries
```{r Load Libraries}
library(tidyverse)
library(readxl)
```

Interaction Plot
```{r Interaction Plot}
detergent <- rep(c("super", "best"), each = 8)
watertemp <- rep(c("cold", "hot"), each = 4)
cleanliness <- c(4, 5, 6, 5, 10, 12, 11, 9, 6, 6, 4, 4, 12, 13, 10, 13)
clothesdf <- data.frame(detergent, watertemp, cleanliness)
clothesdf # "Always double check your data entry" - Professor Wei

model1 <- aov(cleanliness~detergent+watertemp+detergent*watertemp, data = clothesdf)
summary(model1)
interaction.plot(clothesdf$detergent, clothesdf$watertemp, clothesdf$cleanliness, fun = mean)
```
Note: 2.1e-07 = 2.1 x 10^-7

$H_0: \text{There is no interaction between Type and Size}$ 
vs 
$H_a: \text{There is a significant interaction between Type and Size}$
Car Interaction Plot
```{r Car Interaction Plot}
Car_filter <- read_excel("~/Desktop/STAT 301/Week 4/Car filter.xlsx")
model2 <- aov(Noise~Size+Type+Size*Type, data=Car_filter)
summary(model2)
interaction.plot(Car_filter$Type, Car_filter$Type, Car_filter$Noise)
```
f = 30.42; p-value = 0
Reject H0; There is a significant interaction between Type and Size.

standardfilter
```{r standardfilter}
standardfilter <- Car_filter %>%
  filter(Type == "standard")
# standardfilter
```

octelfilter
```{r octelfilter}
octelfilter <- Car_filter %>%
  filter(Type == "octel")
# octelfilter
```

Test Noise by Size
```{r Test Noise by Size}
t.test(Noise~Size, data = standardfilter, alternative = "two.sided")
```
t = -2.5359; p-value = 0.04434
Reject H0; There is significant difference between medium and large among the 
standard filters.

$H_0: \mu_{med} = \mu_{large}$ 
Test Noise by Size Again
```{r Test Noise by Size Again}
t.test(Noise~Size, data = octelfilter, alternative = "two.sided")
```
t = 13.624; p-value = 5.366e-06
Reject H0; there is sig. diff. bt. medium and large among the octel filters

1. A clinician would like to study the effects of two different drugs (drug A 
and drug B) on systolic blood pressure (SBP) for patients. He would also like to 
consider the effects of gender on systolic blood pressure. He randomly chosen 4 
females and assigned with drug A, 4 females with drug B, 4 males with drug A and 
4 males with drug B. He measured the SBP after each patient taking the
medication. Assuming all of the patients had similar baseline SBP. Test the main 
effects of drug and gender, and the interaction between drug and gender
$H_0: \text{There is no significant difference between drug A and B}$
$H_0: \text{There is no significant difference between males and females}$
Drug Types
```{r Drug Types}
gender1 <- rep(c("Female", "Male"), each = 8)
drug <- rep(c("A", "B"), each = 4) # 4 A's and 4 B's
SBP <- c(120, 110, 100, 105, 99, 101, 102, 98, 100, 101, 100, 101, 121, 119, 103, 121)
BPdf <- data.frame(gender1, drug, SBP)
# BPdf

model3 <- aov(SBP~gender1+drug+gender1*drug, data = BPdf)
summary(model3)

```
