---
title: "Homework 6"
author: "Peyton Hall"
date: "02/26/2025"
output:
  pdf_document:
    latex_engine: lualatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load Libraries
```{r}
library(nlme)
library(tidyr)
library(multcomp) # for glht
```


1. A researcher would like to test the effects of two drugs (A and B) on 
treating Breast cancer. The researcher recruited 10 patients and assigned drug A 
randomly to 5 of them and drug B to the other 5. The researcher measured the 
Oncotype DX score to those 10 patients at three times: Immediately after taking 
the drugs, 1 year after the drugs, and 5 years after. The Oncotype DX score of 
15 or below means that no or low recurrence risk, the score of 21-25 means a 
medium risk of recurrence, and the score of 26-100 means a high risk of 
recurrence. Use the 0.05 significance level to test if there is a significant 
mean score difference between drug A and B, and if there is a significant mean 
score difference between the three times, and if there is a significant 
interaction between drug and time.
Drug, Subject, RightAfter, OneYear, FiveYear
      1,       22,         20,      15
      2,       16,         12,      10
A     3,       23,         11,      10
      4,       25,         14,      12
      5,       26,         18,      16
      6,       25,         21,      21
      7,       20,         18,      19
B     8,       35,         27,      26
      9,       39,         35,      36
      10,      27,         27,      25

Question 1 Code
```{r Question 1 Code}
Drug1 <- c(rep("A", 5), rep("B", 5))
Subject <- 1:10
RightAfter <- c(22, 16, 23, 25, 26, 25, 20, 35, 39, 27)
OneYear <- c(20, 12, 11, 14, 18, 21, 18, 27, 35, 27)
FiveYear <- c(15, 10, 10, 12, 16, 21, 19, 26, 36, 25)
question_one_data <- data.frame(Drug1, Subject, RightAfter, OneYear, FiveYear)
question_one_data

# c) use lme (linear mixed effects) and anova()
library(nlme)
library(tidyr)
long_data <- pivot_longer(question_one_data, cols = c("RightAfter", "OneYear", "FiveYear"), 
                          names_to = "Time", values_to = "Score")
# convert Time to a factor to avoid error in glht fucntion
long_data$Time <- factor(long_data$Time, levels = c("RightAfter", "OneYear", "FiveYear"))
# Convert Drug1 to a factor in the long_data DataFrame
long_data$Drug1 <- factor(long_data$Drug1)
# fit linear mixed-effects model using nlme
model1 <- lme(Score~Drug1+Time+Drug1*Time, random=~1+Time|Subject, data= long_data)
anova(model1)

# i)
right_after_data <- long_data[long_data$Time == "RightAfter",]
one_year_data <- long_data[long_data$Time == "OneYear",]
five_year_data <- long_data[long_data$Time == "FiveYear",]
posthoc_rightafter <- lme(Score~Drug1, random =~1|Subject, data = right_after_data)
posthoc_oneyear <- lme(Score~Drug1, random =~1|Subject, data = one_year_data)
posthoc_fiveyear <- lme(Score~Drug1, random =~1|Subject, data = five_year_data)
posthoc_rightafter<-glht(posthoc_rightafter, linfct = mcp (Drug1 ="Tukey"))
posthoc_oneyear<-glht(posthoc_oneyear, linfct = mcp (Drug1 ="Tukey"))
posthoc_fiveyear<-glht(posthoc_fiveyear, linfct = mcp (Drug1 ="Tukey"))
summary(posthoc_rightafter)
summary(posthoc_oneyear)
summary(posthoc_fiveyear)

```

a) What type of design is this?
two-way mixed repeated measure
b) Formulate the null and alternative hypotheses for testing the main effect of 
drug using symbols.
$H_0: \mu_\text{score drug A} = \mu_\text{score drug B}$ vs $H_a: \mu_\text{score drug A} ≠ \mu_\text{score drug B}$
c) What is the F statistic and p-value testing the main effect of drug?
f = 20.13944; p-value = 0.0020
d) What conclusion can you make based on the p-value in b)?
Reject H0; there is evidence to support the claim that there is a significant 
mean score difference between drug A and B.
e) Formulate the null and alternative hypotheses for testing the main effect of 
time using symbols.
$H_0: \mu_\text{Right After} = \mu_\text{One Year} = \mu_\text{Five Year}$ vs $H_a: \text{At least one mean differs}$
f) What is the F statistic and p-value for testing the main effect of time?
f = 31.34159; p-value <.0001
Reject H0; there is evidence to support the claim that there is a significant 
mean score difference between the three times.
g) What conclusion can you make about whether there is a significant mean score 
difference between the three times?
Due to the low p-value of <.0001, there is a significant mean score difference 
between the three times.
h) Is there a significant interaction between time and drug? What is the p-value 
for that?
There is a significant interaction between time and drug. The p-value is 0.0024.
i) If there is a significant interaction, we need to separate the data file. 
Please separate the data by timing:
  a. For the timing of Right After, is there a significant mean score difference 
  between drug A and B? What is that p-value?
  p-value = 0.0781. i.e. > 0.05, so there is no significant mean score 
  difference between Drug A and Drug B immediately after administration.
  b. For the timing of One year after, is there a significant mean score 
  difference between drug A and B? What is that p-value?
  p-value = 0.00182. i.e. < 0.05, so there is a significant mean score 
  difference between Drug A and Drug B one year after administration.
  c. For the timing of Five year After, is there a significant mean score 
  difference between drug A and B? What is that p-value?
  p-value = 6.23e-05 (0.0000623). i.e. < 0.05, so there is a significant mean 
  score difference between Drug A and Drug B five years after administration.




2. A physician would like to see if different doses of two drugs A and B have 
significantly different effect on reducing headache. The physician recruited 10 
patients and randomly assigned drug A to 5 of them and B to the other five. The 
physician prescribed one pill for the low dose and two pills for the high dose 
and asked the patients to take the low dose first and measured their pain scores 
after, and then to take the high dose and measured their pain scores after. The 
higher score, the more pain. Use the 0.05 significance level to test if there is 
the pain scale is significantly different between drug A and B, between the high 
and low doses, and if there is a significant interaction between the type of 
drug and the dose.
    Individuals, LowDose, HighDose
    1,           52,      50
    2,           61,      58
A   3,           59,      51
    4,           37,      34
    5,           49,      41
    6,           43,      33
    7,           32,      30
B   8,           21,      20
    9,           29,      21
    10,          26,      22
    
$H_0: \mu_{\text{Pain Drug A}} = \mu_{\text{Pain Drug B}}$ vs $H_a: \mu_{\text{Pain Drug A}} \neq \mu_{\text{Pain Drug B}}$
$H_0: \mu_{\text{High Dose}} = \mu_{\text{Low Dose}}$ vs $H_a: \mu_{\text{High Dose}} \neq \mu_{\text{Low Dose}}$
Question 2 Code
```{r Question 2 Code}
Drug2 <- c(rep("A", 5), rep("B", 5))
Individuals <- 1:10
LowDose <- c(52, 61, 59, 37, 49, 43, 32, 21, 29, 26)
HighDose <- c(50, 58, 51, 34, 41, 33, 30, 20, 21, 22)
question_two_data <- data.frame(Drug2, Individuals, LowDose, HighDose)
question_two_data

# b) use lme and anova()
# reshape data to long format to model dose effects separately
question_two_long <- reshape(question_two_data, varying = list(c("LowDose", "HighDose")),
                             v.names = "PainScore", times = c("Low", "High"), 
                             timevar = "Dose", direction = "long", new.row.names = 1:20)
# DV = PainScore; IV = Drug2 and Dose
model2<-lme(PainScore~Drug2+Dose+Drug2*Dose, random=~1+Dose|Individuals, data= question_two_long)
anova(model2)

```

a) What type of design is this?
two-way mixed repeated measure
b) What is the F statistic and p-value for comparing the drug A and B?
f = 19.40837; p-value = 0.0023
c) What conclusion can you make based on the p-value in a)?
Reject H0; there is evidence to support the claim that the pain scale is 
significantly different between drug A and B.
d) What is the F statistic and p-value for comparing the high and low dosage?
f = 20.26169; p-value = 0.0020
e) What conclusion can you make based on the p-value in c)?
Reject H0; there is evidence to support the claim that the pain scale is 
significantly different between the high and low doses.
f) What is the F statistic and p-value for the interaction? Is there a 
significant interaction?
f = 0.00844; p-value = 0.9291
There is no significant interaction between the type of drug and the dose.




3. Food scientists would like to test the difference of pasta’s taste while 
cooking under different temperature and different cooking time. They designed 
the experiment to cook 16 plates of pasta, with 4 under high temperature and 
long cooking time, 4 under high temperature and short cooking time, 4 under 
medium temperature and long cooking time, and 4 under medium temperature and 
short cooking time. At the end of the experiment, they measured the taste scores 
of the 16 plates of pasta. The higher score, the better pasta. Use the 0.05 
significance level to compare if there is a significant difference cooking under 
the two temperature levels, and with the two different time, and if there is a 
significant interaction between temperature and time. The data of taste scores 
is here:
                    MedTemp,     HighTemp
Short Cooking Time, 62,59,41,33, 82,96,95,90
Long Cooking Time,  32,31,45,67, 89,76,88,82

Question 3 Code
```{r Question 3 Code}
Time <- rep(c("Short Cooking Time", "Long Cooking Time"), each = 8)
Temperature <- rep(c("MedTemp", "HighTemp"), each = 4)
Scores <- c(62, 59, 41, 33, 82, 96, 95, 90, 32, 31, 45, 67, 89, 76, 88, 82)
data3 <- data.frame(Time, Temperature, Scores)
data3

# c) use aov() function due to it being two-way ANOVA
# DV = taste scores
model3 <- aov(Scores~Time+Temperature+Time*Temperature, data = data3)
summary(model3)
```

a) What type of design is this? (Note this was from one of our previous class)
two-way ANOVA
b) Formulate the null and alternative hypotheses for comparing the mean scores 
between the short and long cooking time
$H_0: \mu_\text{Short Term} = \mu_\text{Long Term}$ vs $H_a: \mu_\text{Short Term} ≠ \mu_\text{Long Term}$
c) What is the test statistic and p-value for a)?
f = 1.039; p-value = 0.328
d) What conclusion can you draw based on the p-value in b)?
Fail to reject H0; there is no evidence to support the claim that there is a 
significant difference in cooking under the two different times.
e) What is the test statistic and p-value for comparing the mean scores between 
the medium and high temperature levels?
f = 48.520; p-value = 1.51e-05 (i.e. 0.0000151)
f) What conclusion can you draw based on the p-value in d)?
There is evidence to support the claim that there is a significant difference 
cooking under the two temperature levels.
g) Is there a significant interaction between cooking time and temperature? What 
is the p-value?
There is no significant interaction between cooking time and temperature because
the p-value of 0.868 is greater than the significance level.




4. The researcher would like to know the effects of pedagogy A and B on 
improving students’ performance in reading. He recruited 8 students. He trained 
all 8 students with A and measured their performance in month 1 and then month 
2. He then trained all of them with B and measured their performance in a 
different test in month 1 and the month 2. The data is below. Use the 0.05 
significance level to test the main effect of pedagogy and main effect of time, 
and the interaction.
Students, Pedagogy A,       Pedagogy B
          Month 1, Month 2, Month 1, Month 2
1,        67,      66,      82,      88
2,        71,      73,      83,      86
3,        72,      75,      88,      84
4,        79,      77,      89,      82
5,        77,      75,      91,      90
6,        65,      72,      90,      89
7,        63,      60,      95,      88
8,        62,      61,      97,      93

Question 4 Code
```{r Question 4 Code}
Students <- 1:8
Pedagogy<-rep(c("Pedagogy A","Pedagogy B"), each =16) # total data points
Pedagogy<-as.factor(Pedagogy)
month<-rep(c("Month 1","Month 2"),each=8) # number of observations in column
Performance<-c(67,71,72,79,77,65,63,62, 66,73,75,77,75,72,60,61,
               82,83,88,89,91,90,95,97, 88,86,84,82,90,89,88,93)
myperformance<-data.frame(Students,Pedagogy,month,Performance)
myperformance
  
# b) 
model4<-lme(Performance~Pedagogy+month+Pedagogy:month, random = ~ 1|Students, data=myperformance)
anova(model4)

```

a) What is the design?
two-way fixed repeated measure
b) What is the test statistic and p-value for comparing the pedagogies?
c) What conclusion can you make with the p value in b)?
f = 91.278; p-value = <.0001
Reject H0; there is evidence to support the claim that the mean performances 
between pedagogy a and pedagogy b differ.
d) What is the test statistic and p-value for comparing the timing?
f = 0.146; p-value = 0.7062
e) What conclusion can you make with the p-value in d)?
Fail to reject H0; there is not enough evidence to support the claim that 
month 1 and month 2 performances differ.
f) What is the p-value for testing the interaction? Is the interaction 
significant?
p-value = 0.5726
There is not enough evidence to support the claim that there is a significant 
interaction.




5. An online retailer wants to get the best from its employees, as well as 
improving their working experience. Currently, employees in the retailer’s order 
fulfilment center are not provided with any kind of entertainment whilst they 
work (e.g., no background music). However, the retailer wants to know whether 
providing music, which a few employees have requested, would lead to greater 
productivity. He recruited a sample of 15 volunteers and split them into three 
treatment groups: a) a control group that didn’t listen to music; b) a treatment 
group who listened to music but had no choice of what they listened to; and c) a 
second treatment group who listened to music and had a choice of what they 
listened to. The data of their productivities is provided on D2L. The data is 
not normally distributed. Use the 0.05 significance level to test that if there 
is any significant difference in their productivities among the three groups.

Question 5 Code
```{r Question 5 Code}
library(readxl)
EmployeeMusic <- read_excel("~/Desktop/STAT 301/Week 6/EmployeeMusic.xlsx")

# DV = Productivity, IV = Music
kruskal.test(Productivity~Music, data = EmployeeMusic)
```

Step 1: State the null and alternative hypotheses in symbols
$H_0: \eta_\text{ProductivityA} = \eta_\text{ProductivityB} = \eta_\text{ProductivityC}$ vs $H_a: \text{At least two medians are different}$
Step 2: Choose Kruskal-Wallis rank sum test
Step 3: What is the test statistic and p-value from R?
chi-squared = 0.093644; p-value = 0.9543
Step 4: State your decision to H0 and explain the decision in the context
Fail to reject H0; there is not enough evidence to support the claim that 
productivities among the three groups are not equivalent.

