---
title: "Worksheet 6"
author: "Peyton Hall"
date: "02/21/2025"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(nlme)
library(multcomp)
```

Activity 1
We would like to study the effect of pizza consumption and the timing of
eating pizza on mood. The study recruited 16 volunteers, and randomly
assigned 8 of them to eat Pop Johns’ Pizza, and the other 8 eat Domino’s Pizza.
All 16 volunteers had Pizza for both breakfast and dinner. The researchers
recorded their mood scores 3 hours after they had Pizza. The data is on the
worksheet.
Perform an appropriate analysis to test the main effect of the type of Pizza, the
main effect of timing, and the interaction between type of Pizza and timing

Individual, Type of Pizza, After Breakfast, After dinner
1           Pop John's,    52,              77
2                          59,              69
3                          66,              68
4                          43,              71
5           Dominos,       55,              82
6                          61,              69
7                          62,              80
8                          39,              59

```{r}
individuals <- 1:8
pizza <- rep(c("Pop Johns", "Dominos"), each = 4)
meals <- rep(c("After breakfast", "After dinner"), each = 8)
moodscore <- c(52,59,66,43, 55,61,62,39, 77,69,68,71, 82,69,80,59)
mydf <- data.frame(individuals, pizza, meals, moodscore)
model2 <- lme(moodscore~pizza+meals+pizza*meals, random = ~1+meals|individuals, data=mydf)
anova(model2)
```

a) What are the f statistic and p-value for testing the main effect of type of
Pizza?
f = 0.0142; p-value = 0.9089

b) What conclusion can you make about the main effect of Pizza?
$H_0: \mu_{PJ} = \mu_{D}$ vs $H_a: \mu_{PJ} ≠ \mu_{D}$
Fail to reject H0; there is no significant difference in mood score between Pop
John's and Dominos pizza.

c) What are the F statistic and p-value for testing the main effect of 
timing/meal?
f = 22.2650; p-value = 0.0033

d) What conclusion can you make about the main effect of timing/meal?
$H_0: \mu_{AB} = \mu_{AD}$ vs $H_a: \mu_{AB} ≠ \mu_{AD}$
Reject H0; there is significant mood score difference between breakfast and
dinner

e) What is the p-value for the interaction term?
p-value = 0.7936
Fail to reject H0; there is no significant interaction between pizza and meal.


2. Does physical exercise alleviate depression? We find some depressed people
and check that they are all equivalently depressed to begin with. Then we
allocate each person randomly to one of the three groups: no exercise; 20
minutes of jogging per day; or 60 minutes of jogging per day. At the end of a
month, we ask each participant to rate how depressed they feel on a Likert
scale from 1 (totally miserable) through 100 (ecstatically happy). Here is the
data. Use 0.05 significance level to test the claim that the depression level
(score) are significantly different between the three exercise groups.
No exercise, 23, 26, 32, 37, 29
20-min exercise, 22, 27, 39, 35, 46
60-min exercise, 59, 66, 38, 49, 56

```{r}
exercise <- rep(c("No exercise", "20-min exercise", "60-min exercise"), each = 5)
depscore <- c(23, 26, 32, 37, 29, 22, 27, 39, 35, 46, 59, 66, 38, 49, 56)
depdf <- data.frame(exercise, depscore)
kruskal.test(depscore~exercise, data = depdf)
pairwise.wilcox.test(depscore, exercise, p.adj="none")
```


Step 1: State the null and alternative hypotheses
$H_0: \text{μ Depression no exercise = μ Depression twenty min = μ Depression sixty min}$ vs $H_a: \text{At least two μ are different}$

Step 2: What test would you choose?
Kruskal-Wallis rank sum test

Step 3: What is the test statistic and p-value?
t = 8.34; p-value = 0.01545

Step 4: What is your decision?
Reject H0; there is a sig. diff. bt. at least 2 types of exercises.

Step 5: What can you conclude from post-hoc (follow up) test?
The significant difference is 
bt. 20 min and 60 min
bt. none and 60 min




