---
title: "Homework 3"
author: "Peyton Hall"
date: "02/07/2025"
output:
  pdf_document:
    latex_engine: lualatex
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load Necessary Libraries
```{r Load Necessary Libraries}
library(readxl)
library(tidyverse)
```


1. Dr. Brown would like to compare his students’ performance after he taught 
three classes with three different pedagogies. He used his students’ final exam 
scores as the measurements. He took a sample of 12 students’ final scores from 
pedagogy A, 12 from pedagogy B and 12 from C. Use the 0.05 significance level to 
test the claim that there is a significant difference between the mean final 
scores of the three pedagogies. The data is on D2L, named as Student final 
scores.
a) Step 1: Formulate the null and alternative hypotheses in symbols
$H_0: μ_A = μ_B = μ_C$ vs $H_a: \text{At least one μ differs}$
Step 2: Choose the one-way ANOVA test
b) Step 3: Find the correct test statistic and p-value for testing the means 
from R output
c) Step 4: Make a decision based on the significance level of 0.05; and explain 
your decision in the context.
d) If there are significant differences, perform a pairwise comparison to 
determine where the differences lie.
Question 1 Code
```{r Question 1 Code}
Student_final_scores <- read_excel("~/Desktop/STAT 301/Week 3/Student final scores.XLSX")

# Step 3
anova_result1 <- aov(Scores ~ Pedagogy, data = Student_final_scores)
summary(anova_result1)

# Step 4 d)
TukeyHSD(anova_result1) # pairwise comparison test

```
f = 5.926; p-value = 0.00632
Reject H0; there is evidence to support the claim that there is a significant 
difference between the mean final scores of the three pedagogies.
In the pairwise comparison test, it is found that the mean of B significantly 
differs from that of A, the mean of C significantly differs from that of A, and
the mean of C does not significantly differ from that of B.


2. The data CO2 on d2l recorded the CO2 emission from the burning of fossil 
fuels (metric tonnes of CO2 per person) (Source: https://cdiac.ess-dive.lbl.gov/ 
). The data recorded the CO2 emission per person from three countries: USA, 
Russia and India. Use the 0.05 significance level to test the claim that the 
average CO2 emission per person are significantly different between the three 
counties. Note, the format of the data is wide format. You will need to use the 
gather()OR the pivot_longer()function in R to convert it to a long format with 
one independent variable and one dependent variable so the format is analysis 
ready.
a) Step 1: Formulate the null and alternative hypotheses in symbols
$H_0: μ_{USA} = μ_{Russia} = μ_{India}$ vs $H_a: \text{At least one μ differs}$
Step 2: Choose the one-way ANOVA test
b) Step 3: Find the correct test statistic and p-value for testing the means 
from R output
c) Step 4: Make a decision based on the significance level of 0.05; and explain 
your decision in the context.
d) If you discover the significant difference, Where the differences lie?
Question 2 Code
```{r Question 2 Code}
CO2 <- read_excel("~/Desktop/STAT 301/Week 3/CO2.XLSX")
# Step 3
CO2_long <- pivot_longer(
  CO2,
  cols = c(`United States`, Russia, India),
  names_to = "Country",
  values_to = "Emissions"
)
anova_result2 <- aov(Emissions ~ Country, data = CO2_long)
summary(anova_result2)

# 2 d)
TukeyHSD(anova_result2) # pairwise comparison test

```
f = 1524; p-value = <2e-16
Reject H0; there is evidence to support the claim that the average CO2 emission 
per person are significantly different between the three counties.
In the pairwise comparison test, it is found that the mean CO2 emissions per 
person significantly differ between each pair of countries analyzed. 
Specifically, the mean emissions from the United States are significantly higher 
than those from India by approximately 18.2 metric tonnes and also higher than 
those from Russia by about 8.2 metric tonnes. Additionally, the mean emissions 
from Russia are significantly higher than those from India by approximately 10.0 
metric tonnes. These results indicate substantial differences in CO2 emissions 
per capita among the three countries.


3. A clinical research was conducted to compare three treatments (drug A, B and 
C) on treating depression. The researchers recruited 24 volunteers and randomly 
assigned 8 to take drug A, 8 to take B, and 8 to take C. The baseline 
characteristics of the 24 volunteers are the same. The researchers measured the 
depression scores after the volunteers took the medications. The data is listed 
in the table below. Suppose the data follow normal distribution. The lower 
depression score, the more effective the drug is. Use the 0.05 significance 
level to test the claim that the average depression scores after taking the 
drugs are significantly different.
Drug A, 51, 45, 33, 41, 36, 38, 39, 33
Drug B, 23, 31, 23, 20, 19, 26, 28, 17
Drug C, 22, 18, 29, 32, 41, 20, 16, 23
a) Enter the data in R with the correct format (one IV and one DV)
b) Step 1: Formulate the null and alternative hypotheses in symbols
$H_0: μ_A = μ_B = μ_C$ vs $H_a: \text{At least one μ differs}$
Step 2: Choose the one-way ANOVA test
c) Step 3: Find the correct test statistic and p-value for testing the means 
from R output
d) Step 4: Make a decision based on the significance level of 0.05; and explain 
your decision in the context.
e) Which two drugs had significantly different average depression scores?
Question 3 Code
```{r Question 3 Code}
# a) Enter data
Drug <- rep(c("A", "B", "C"), each = 8)
DepressionScore <- c(51, 45, 33, 41, 36, 38, 39, 33,  # Drug A scores
                     23, 31, 23, 20, 19, 26, 28, 17,  # Drug B scores
                     22, 18, 29, 32, 41, 20, 16, 23)  # Drug C scores
depressiondf <- data.frame(Drug, DepressionScore)

# Step 3, use aov function to perform analysis of variance
anova_result3 <- aov(DepressionScore ~ Drug, data = depressiondf)
summary(anova_result3)

# 3 e)
TukeyHSD(anova_result3) # pairwise comparison test

```
f = 14.44; p-value = 0.000113
Reject H0; there is evidence to support the claim that the average depression 
scores after taking the drugs are significantly different.
In the pairwise comparison test, it is found that the mean depression scores for 
Drug B and Drug C are significantly lower than those for Drug A. Specifically, 
the mean depression score for Drug B is significantly lower than that for Drug A 
by approximately 16.1 points, and the mean score for Drug C is also lower than 
Drug A by about 14.4 points. However, there is no significant difference between 
the depression scores for Drug B and Drug C.


4. A representative from a seed-producing company has been on a seed-collecting 
trip in the hope of finding improved varieties of millet. She wants to discover 
whether the yields from different local seed sources are significantly 
different. An experiment is set up with five different seed sources (A, B, C, D 
and E). For each seed source, she planted seven replications of plots. She 
measured the yield from each plot at the end of the growing season. She is 
interested in testing the main effect of seed sources with the significance 
level of 0.05. The data is as following. Perform an appropriate test to answer 
the following questions:
  A,   B,   C,   D,   E
1.4, 1.5, 0.3, 1.5, 1.1
1.2, 1.8, 0.5, 1.9, 1.5
1.1, 1.6, 0.8, 1.2, 0.7
1.0, 1.3, 0.2, 1.6, 1.4
0.9, 1.2, 0.6, 1.8, 1.2
1.8, 1.9, 0.5, 1.5, 1.3
1.2, 1.7, 0.3, 0.6, 1.7

a) Enter the data in R with the correct format.
b) Step 1: State the null and alternative hypotheses
$H_0: μ_A = μ_B = μ_C = μ_D = μ_E$ vs $H_a: \text{At least one μ differs}$
Step 2: Choose the one-way ANOVA test
c) Step 3: Find the test statistic and p-value for testing the significant 
difference in mean
d) Step 4: What conclusion can you draw based on the p-value and the 
significance level of 0.05?
e) Step 5: Which two seed sources are significantly different from each other 
based on the significance level of 0.05?
Question 4 Code
```{r Question 4 Code}
# a) Enter data in R
A <- c(1.4, 1.2, 1.1, 1.0, 0.9, 1.8, 1.2)
B <- c(1.5, 1.8, 1.6, 1.3, 1.2, 1.9, 1.7)
C <- c(0.3, 0.5, 0.8, 0.2, 0.6, 0.5, 0.3)
D <- c(1.5, 1.9, 1.2, 1.6, 1.8, 1.5, 0.6)
E <- c(1.1, 1.5, 0.7, 1.4, 1.2, 1.3, 1.7)
yields <- data.frame(A, B, C, D, E)
yields

# Step 3: Find test statistic
yields <- data.frame(
  value = c(A, B, C, D, E),
  seed_source = factor(rep(c("A", "B", "C", "D", "E"), each = 7))
) # data frame in long format created

anova_result4 <- aov(value ~ seed_source, data = yields)
summary(anova_result4)

# Step 5
TukeyHSD(anova_result4) # pairwise comparison test
```
f = 13.49; p-value = 2.11e-06
Reject H0; there is evidence to support the claim that the yields from different 
local seed sources are significantly different.
In the pairwise comparison test, it is found that the mean of C significantly 
differs from that of A, B, D, and E.


5. Suppose that we have a different data for the chest deceleration for three 
different size categories (small, midsize and large) of cars. In the sample, we 
have a total of 21 cars. We still want to use the significance level of 0.05 and 
test the claim that the different size categories have the different chest 
deceleration in the standard crash test. Suppose the equal variance assumption 
is satisfied. Part of the output from SPSS is listed below. Using the 
information from the SPSS output, calculate the MS_B, MS_W, df_B, df_W, and the 
F statistic.

Sources,        Sum of Squares, Df, Mean Square, F Statistic, P-Value
Between Groups, 200.857,        ?,  ?,           ?,           0.061
Within Groups,  549.714,        ?,  ?,           ?,           0.061

Show your calculations:

MS_B = Mean Square Between Groups
MS_W = Mean Square Within Groups
df_b = degrees of freedom between groups
df_w = degrees of freedom within groups

df_b = # of groups - 1 = k - 1 = 3 - 1 = 2

df_w = total observations - # of groups = n - k = 21 - 3 = 18

MS_B = (SS_B) / (df_B) = 200.857 / 2 = 100.4285

MS_W = (SS_W) / (df_W) = 549.714 / 18 = 30.5396

F = (MS_B) / (MS_W) = 100.4285 / 30.5396 = 3.28846809