---
title: "Worksheet 3"
author: "Peyton Hall"
date: "01/31/2025"
output:
  pdf_document:
    latex_engine: lualatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
```

```{r}
drugtype <- rep(c("drug1", "drug2", "drug3"), c(5, 7, 7))
relieftime <- c(7.3, 8.2, 10.1, 6, 9.5, 7.1, 10.6, 11.2, 9, 8.5, 10.9, 7.8, 5.8, 
                6.5, 8.8, 4.9, 7.9, 8.5, 5.2)
myrelief <- data.frame(drugtype, relieftime)
myrelief

model1 <- aov(relieftime~drugtype, data = myrelief)
summary(model1)
```
f = 4.188; p-value = 0.0345
Reject H0; there is no evidence to support that the average relief time for the 
three drugs are indentical.

```{r}
myrelief %>%
  group_by(drugtype) %>%
  summarize(avetime = mean(relieftime), setime = sd(relieftime)/
              sqrt(length(relieftime)), 
            tstar=qt(1-0.05/2, length(relieftime) - 1))%>%
  ggplot(aes(x=drugtype, y=avetime)) + geom_point()+
  geom_errorbar(aes(ymin=avetime -setime*tstar, ymax = avetime+setime*tstar))+
  ylim(c(0,11))
```

There is various flickering light in our environment. For instance, light from
computer screens and fluorescent bulb. If the frequency of the flicker is below 
a certain threshold, the flicker can be detected by the eye. Different people 
have different flicker "threshold" frequencies (known as the critical flicker 
frequency, or CFF). Knowing the critical threshold frequency below which flicker
is detected can be important for product manufacturing as well as tests for 
ocular disease. Do people with different eye color have different threshold 
flicker sensitivity? A 1973 study ("The Effect of Iris Color on Critical Flicker 
Frequency", Journal of General Psychology (1973)) obtained the following data 
(see worksheet) from a random sample of 19 subjects. Do these data suggest that
people with different eye color have different threshold sensitivity to 
flickering light?
Worksheet 3 Number 1
```{r Worksheet 3 Number 1}
eyecolor <- rep(c("Brown", "Green", "Blue"), c(8,5,6))
CFF <- c(26.8, 27.9, 23, 25, 26, 24.8, 25.7, 24.5, 26.4, 24.2, 28, 26.9, 29.1, 
         25.7, 27.2, 29.9, 28.5, 29.4, 29.4)
flickerdf <- data.frame(eyecolor, CFF)

model2 <- aov(CFF~eyecolor, data=flickerdf)
summary(model2)
TukeyHSD(model2) # pairwise comparison test


# Step 1: formulate the null and alternative hypotheses using symbols
# H_0: μ_Brown = μ_Green = μ_Blue
# H_a: at least 2 means are different

# Step 2: choose one-way ANOVA test.

# Step 3: Find the statistic and the p-value for testing the means from R output

# Step 4: Make a decision based on significance level of 0.05; and explain your
# decision in the context.

# Step 5: If there is any significant differences among the means, perform a 
# pairwise comparison to determine where the differences lie.
```
f = 5.442; p-value = 0.0157
Reject H0; there is evidence to support that people with diff eye color have 
diff threshold sensitivity to flickering lights.

2. Referring to the data "James Intrusive Memory Activity", perform the 
following one-way ANOVA tests by answering the questions below. The authors 
compared the frequency of intrusive memories under four conditions: No task(1); 
Reactivation plus Tetris (2); Reactivation only (3); and Tetris only (4). The 
independent variable is the first column of the data '"Condition"
```{r}
James_Intrusive_Memories_Activity <- read_excel("~/Desktop/STAT 301/Week 3/James Intrusive Memories Activity.xlsx")
James_Intrusive_Memories_Activity
# a) Test to see if the mean frequency of intrusive memories are significantly 
# different between the four conditions at day O, using the variable 
# Day_Zero_Number_of_Intrusions with the 0.05 significance level.
model4 <- aov(Day_Zero_Number_of_Intrusions~Condition, 
              data=James_Intrusive_Memories_Activity)
summary(model4)

# b) Test to see if the mean frequency of intrusive memories are significantly 
# different between the four conditions for the seven days following the 
# experiment, using the variable Days_One_to_Seven_Number_of_Intrusions, with 
# the significance level of 0.05.
model4 <- aov(Days_One_to_Seven_Number_of_Intrusions~Condition, 
              data=James_Intrusive_Memories_Activity)
summary(model4)

# C) continue with b), if you observe any significant difference, perform a 
# multiple comparison to determine where the significant differences lie.
TukeyHSD(model4) # pairwise comparison test
```
a)
f = 0.163; p-value = 0.921
Fail to reject H0; there is not enough evidence to support that the mean 
frequency of intrusive memories are significantly different between the four 
conditions at day O

b)
f = 3.795; p-value = 0.0141
Reject H0; there is evidence to support that the mean frequency of intrusive 
memories are significantly different between the four conditions for the seven 
days following the experiment




