---
title: "Homework 5"
author: "Peyton Hall"
date: "02/19/2025"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(nlme)
library(multcomp)
```


1. A researcher has randomly selected a sample of 8 adults to participate in a 
memory study. Each adult is administered a treatment and then given a memory 
test. The memory test scores are standardized. A higher score indicates a better 
memory score. The data is shown below. Use the 0.05 significance level to test 
that there is a significant mean difference between the test scores of taking 
the three treatments, by answering the following questions.
Subject, Control, DrugA, DrugB
1,       5,       9,     6
2,       4,       9,     4
3,       5,       10,    7
4,       6,       11,    9
5,       7,       12,    10
6,       3,       10,    3
7,       4,       8,     5
8,       5,       9,     4
a) Formulate the null and alternative hypotheses to test the differences of the 
treatments.
$H_0: μ_\text{Control} = μ_\text{DrugA} = μ_\text{DrugB}$ vs $H_a: \text{At least one μ differs}$
b) What are the F statistic and the p-value? What do you conclude based on the 
significance level of 0.05?
c) If there is a significant difference between treatments, which two are 
significantly different?

1
```{r 1}
control <- c(5, 4, 5, 6, 7, 3, 4, 5)
drugA <- c(9, 9, 10, 11, 12, 10, 8, 9)
drugB <- c(6, 4, 7, 9, 10, 3, 5, 4)

data1 <- data.frame(
  Subject = rep(1:8, 3),
  Score = c(control, drugA, drugB),
  Treatment = factor(rep(c("Control", "DrugA", "DrugB"), each = 8))
)

# use lme due to one way repeated measure 
result1 <- lme(Score~Treatment, random = ~1 | Subject, data = data1)
anova(result1)

# c)
data1$Subject <- as.factor(data1$Subject)
posthoc1<-glht(result1, linfct=mcp(Treatment="Tukey"))
summary(posthoc1)

```
b)
f = 50.6185; p-value = <.0001
Reject H0; there is evidence to support the claim that there is a significant 
mean difference between the test scores of taking the three treatments.
c)
There is a sig. diff. bt. DrugA-Control due to the low p-value.
There is a sig. diff. bt. DrugB-DrugA due to the low p-value.


2. A therapist interested in helping individuals with Ophidiophobia. In order to 
do this, the therapist is going to slowly desensitize the patients to snakes. He 
needs to find a starting point, so he presented a sample of 8 ophidiophobics 
with snake related stimuli and measured their physiologic responses (1-20). Use 
the 0.05 significance level to test the claim that there is a significant mean 
physiologic score difference reacting to the three different stimuli.
Subjects, Picture of a snake, Plastic toy snake, Rubber toy snake
1,        7,                  12,                14
2,        5,                  8,                 13
3,        3,                  10,                12
4,        8,                  12,                14
5,        2,                  7,                 12
6,        4,                  4,                 10
7,        3,                  16,                8
8,        4,                  9,                 13
a) What are the null and alternative hypotheses
$H_0: μ_\text{Picture} = μ_\text{Plastic} = μ_\text{Rubber}$ vs $H_a: \text{At least one μ differs}$
b) What are the test statistic and p-value?
c) What decision can you make in the context?
d) If there is a significant difference, which two?

2
```{r 2}
picture <- c(7, 5, 3, 8, 2, 4, 3, 4)
plastic <- c(12, 8, 10, 12, 7, 4, 16, 9)
rubber <- c(14, 13, 12, 14, 12, 10, 8, 13)

data2 <- data.frame(
  Subject = rep(1:8, 3),
  Score = c(picture, plastic, rubber),
  Treatment = factor(rep(c("Picture", "Plastic", "Rubber"), each = 8))
)

# use lme due to one way repeated measure 
result2 <- lme(Score~Treatment, random = ~1 | Subject, data = data2)
anova(result2)

# d)
data2$Subject <- as.factor(data2$Subject)
posthoc2<-glht(result2, linfct=mcp(Treatment="Tukey")) # pairwise comparison
summary(posthoc2)

```
b) - c)
t = 19.36576; p-value = 1e-04 (.0001)
Reject H0; there is evidence to support the claim that there is a significant 
mean physiologic score difference reacting to the three different stimuli.
d)
There is a sig. diff. bt. Plastic-Picture due to the low p-value.
There is a sig. diff. bt. Rubber-Picture due to the low p-value.


3. Consider an experiment on testing the efficiency of two drugs on lowering 
heart rate for heart disease patients. Eight patients are administered three 
drugs (one of them is the placebo control) at three different years and heart 
rate measured every five minutes for a total of four measurements. Suppose that 
the heart rates did not change significantly through the three years. That means 
there is no year effect on the experiment. Perform an appropriate test to 
examine the main effect of the drugs on lowering heart rate, the main effect of 
time and the interaction between drug and time. The data is given below:
Patient ID,         Drug AX23,                 Drug BWW9,                 Control
            5min, 10min, 15min, 20min, 5min, 10min, 15min, 20min, 5min, 10min, 15min, 20min
1,          82,   86,    81,    77,    85,   66,    83,    80,    69,   73,    72,    74
2,          78,   83,    88,    81,    82,   86,    80,    64,    66,   62,    67,    73
3,          81,   82,    81,    75,    71,   78,    70,    75,    84,   90,    88,    87
4,          82,   83,    83,    89,    83,   88,    79,    81,    80,   81,    77,    72
5,          86,   79,    77,    86,    86,   85,    76,    76,    72,   72,    69,    70
6,          84,   83,    84,    77,    86,   82,    83,    80,    65,   62,    65,    61
7,          82,   73,    78,    70,    79,   83,    80,    81,    75,   69,    69,    68
8,          89,   85,    76,    80,    83,   84,    78,    81,    71,   70,    65,    65
a) Formulate the null and alternative hypotheses for testing the main effect of 
time.
$H_0: μ_\text{5min} = μ_\text{10min} = μ_\text{15min} = μ_\text{20min}$ vs $H_a: \text{At least one μ differs}$
b) What are the F-statistic and p-value for testing the main effect of time?
c) What decision can you make for the main effect of time?
d) Formulate the null and alternative hypotheses for testing the main effect of 
drug.
$H_0: μ_\text{AX23} = μ_\text{BWW9} = μ_\text{Control}$ vs $H_a: \text{At least one μ differs}$
e) What is the F and p-value for testing the main effect of drug?
f) What decision can you make for the main effect of drug?
g) What are the F-value and p-value for the interaction term? What can you 
conclude about the interaction term (is it significant)?
h) If there are significant differences in the heart rate after taking different 
drugs, which two of drugs are significantly different?

3
```{r 3}
# b) - c)
time_5min <- c(82, 78, 81, 82, 86, 84, 82, 89, 85, 82, 71, 83, 86, 86, 79, 83, 69, 66, 84, 80, 72, 65, 75, 71)
time_10min <- c(86, 83, 82, 83, 79, 83, 73, 85, 66, 86, 78, 88, 85, 82, 83, 84, 73, 62, 90, 81, 72, 62, 69, 70)
time_15min <- c(81, 88, 81, 83, 77, 84, 78, 76, 83, 80, 70, 79, 76, 83, 80, 78, 72, 67, 88, 77, 69, 65, 69, 65)
time_20min <- c(77, 81, 75, 89, 86, 77, 70, 80, 80, 64, 75, 81, 76, 80, 81, 81, 74, 73, 87, 72, 70, 61, 68, 65)

data3 <- data.frame(
  Subject = factor(rep(1:8, each = 12)),
  Time = factor(rep(c("5min", "10min", "15min", "20min"), each = 6, times = 2)),
  HeartRate = c(time_5min, time_10min, time_15min, time_20min),
  Drug = factor(rep(c("AX23", "BWW9", "Control"), each = 32))
)

# fit linear mixed-effects model
model1 <- lme(HeartRate~Time, random = ~1 | Subject, data = data3)
anova(model1) # examine the main effect of time

# e) - f)
AX23 <- c(82, 86, 81, 77, 78, 83, 88, 81, 81, 82, 81, 75, 82, 83, 83, 89, 86, 79, 77, 86, 84, 83, 84, 77, 82, 73, 78, 70, 89, 85, 76, 80)
BWW9 <- c(85, 66, 83, 80, 82, 86, 80, 64, 71, 78, 70, 75, 83, 88, 79, 81, 86, 85, 76, 76, 86, 82, 83, 80, 79, 83, 80, 81, 83, 84, 78, 81)
Control <- c(69, 73, 72, 74, 66, 62, 67, 73, 84, 90, 88, 87, 80, 81, 77, 72, 72, 72, 69, 70, 65, 62, 65, 61, 75, 69, 69, 68, 71, 70, 65, 65)

data4 <- data.frame(
  Subject = factor(rep(1:8, each = 12)),
  Time = factor(rep(c("5min", "10min", "15min", "20min"), each = 3, times = 2)),
  HeartRate = c(AX23, BWW9, Control),
  Drug = factor(rep(c("AX23", "BWW9", "Control"), each = 32))
)

# fit linear mixed-effects model
model2 <- lme(HeartRate~Drug, random = ~1 | Subject, data = data4)
anova(model2) # examine the main effect of drug

# g)
data5 <- data.frame(
  Subject = factor(rep(1:8, each = 12)),
  Time = factor(rep(c("5min", "10min", "15min", "20min"), times = 8)),
  Drug = factor(rep(c("AX23", "BWW9", "Control"), each = 4, times = 8)),
  HeartRate = c(time_5min, time_10min, time_15min, time_20min)
)
# fit linear mixed-effects model
model3 <- lme(HeartRate~Time * Drug, random = ~1 | Subject, data = data5)
anova(model3)

# h)
posthoc_drug <- glht(model, linfct = mcp(Drug = "Tukey"))
summary(posthoc_drug)

```
b) - c)
f = 8.06; p-value = 1e-04 (0.0001)
Reject H0; there is evidence to support the claim that the means of heart rate 
measurements differ across different time points.
e) - f)
f = 10.125; p-value = 1e-04 (0.0001)
Reject H0; there is evidence to support the claim that a significant difference 
in heart rate effects between the drugs exists.
g) 
f = 1.202; p-value = 0.3144
The interaction term is not significant. This indicates that the effect of the 
drugs does not change significantly over time.
h) 
There is a sig. diff. bt. Control-AX23 due to the low p-value.
There is a sig. diff. bt. Control-BWW9 due to the low p-value.


4. Do a two-way repeated-measures ANOVA for the following data. We measured 
reading of the altimeters at two different temperatures (low and high) and 
measured the reading at four positions of altimeters (high, low, right, left). 
The DV is accuracy of altimeter reading. High scores show more accurate reading 
than low scores. Perform your analysis and answer the following questions to 
test the claim that there is a significant mean readings between the positions 
and between the temperature levels, and that there is a significant interaction
between temperature and position.
Altimeter,    Low temperature,        High temperature
           High, Low, Left, Right, High, Low, Left, Right
1,         84,   45,  47,   90,    80,   49,  45,   73
2,         69,   28,  12,   68,    33,   43,  18,   46
3,         99,   40,  35,   71,    36,   22,  56,   38
4,         70,   35,  29,   59,    51,   20,  42,   50
5,         91,   18,  19,   85,    37,   44,  38,   39
a) Formulate the null and alternative hypotheses for testing the main effect of 
temperature on the accuracy of reading.
$H_0: μ_\text{Low temperature} = μ_\text{High temperature}$ vs $H_a: μ_\text{Low temperature} ≠ μ_\text{High temperature}$
b) What is the F value and p-value for testing the main effect of temperature?
c) What decision can you make for that main effect of temperature?
d) What is the F and p-value for testing the main effect of position?
e) What decision can you make for the main effect of position?
f) What is the F and p-value for testing the interaction? What can you conclude 
about the interaction term (is it significant)?
g) If there is a significant interaction, separate the data by position and test 
whether there is a significant difference between the low and high temperature 
at each position. (list your t-statistic, p-value and conclusion for each 
position)

4
```{r 4}
# DV = alt reading
# IV1 = temp
# IV2 = 4 positions
# subject = subject #

# b)
subject <- factor(rep(1:5, each = 8))
temperature <- factor(rep(c("Low", "High"), each = 4, times = 5))
position <- rep(c("High", "Low", "Left", "Right"), times = 5, each = 2)
reading <- c(84, 45, 47, 90, 80, 49, 45, 73, 69, 28, 12, 68, 33, 43, 18, 46,
             99, 40, 35, 71, 36, 22, 56, 38, 70, 35, 29, 59, 51, 20, 42, 50,
             91, 18, 19, 85, 37, 44, 38, 39)
data6 <- data.frame(Subject = subject, Temperature = temperature, 
                    Position = position, Reading = reading)
# fit linear mixed-effects model
model4 <- lme(Reading~Temperature, random = ~1 | Subject, data = data6)
anova(model4) # main effect of temperature

# d)
model_position <- lme(Reading~Position, random = ~1 | Subject, data = data6)
anova(model_position)

# f)
model_interaction <- aov(Reading~Temperature * Position + Error(Subject), data = data6)
summary(model_interaction)


```
b) - c)
f = 2.93429; p-value = 0.0958
Fail to reject H0; there is no evidence to support the claim that there is a 
mean difference in the accuracy of altimeter readings between the temperature 
levels.
d) - e)
f = 1.10597; p-value = 0.3611
There is no evidence of a significant difference among the positions.
f) 
f = 0.256; p-value = 0.776
The interaction term is not significant. The effect of temperature on altimeter 
accuracy does not change across positions.


