---
title: "Homework1"
author: "Peyton Hall"
date: "01/22/2025"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Necessary Libraries
```{r Load Necessary Libraries}
library(dplyr)
library(ggplot2)
```


1. An experiment is designed to determine how speaker size affects loudness. The 
   researcher measures loudness of 18 speakers randomly: 6 speakers have small 
   diameters, 6 have medium diameters, and 6 have larger diameters. The 
   researcher records the loudness (in decibels) of the speakers playing the 
   same sound in the table below, and compared the three types of speakers.

   Small diameter speaker loudness, 15, 21, 16, 14, 20, 22
   Medium diameter speaker loudness, 32, 41, 46, 39, 52, 43
   Large diameter speaker loudness, 52, 50, 61, 69, 43, 58

a) Identify the independent variable (IV)
   IV = Speaker Size
b) Identify whether the independent variable is categorical or numeric.
   IV = Ordinal Categorical Variable (i.e. ordered)
c) Identify the dependent variable (DV)
   DV = loudness (in decibels)
d) Identify whether the DV is categorical or numeric
   DV = Discrete Numeric Variable (i.e. integers only)
e) What can be the confounding variables? (list one possible variable)
   One possible confounding variable could be the acoustics of the room.

Question 1 Parts f) and g)
```{r Question 1 Parts f and g}
# f) Enter the data in R as a data frame with one dependent and one independent 
#    variable

# each = 6 ensures each speaker sizes' loudness is repeated six times.
# This matches the six loudness data points to each size.
# rep() function replicates values in a vector
IV_Size <- rep(c("Small", "Medium", "Large"), each = 6)
DV_Loudness <- c(15, 21, 16, 14, 20, 22, 32, 41, 46, 39,
              52, 43, 52, 50, 61, 69, 43, 58)

# create data frame using the vectors
speaker_data <- data.frame(IV_Size, DV_Loudness)
speaker_data

# g) Use pipeline to calculate the mean and standard deviation of loudness for 
#    each type of speaker. List the mean and standard deviation for each speaker
#    here:

speaker_stats <- speaker_data %>%
  group_by(IV_Size) %>%
  summarize(Mean_Loudness = mean(DV_Loudness), SD_Loudness = sd(DV_Loudness))
speaker_stats
```

2. Dr. Optimist believes he has discovered a “smarts” pill will increase IQ. He 
   first measures the IQ of 20 volunteered college juniors and then randomly 
   assigns them to one of the two groups: 10 of them take the smarts pill and 10 
   take a placebo (sugar pill). He then measures the IQ of all the volunteers. 
   He would like to compare the mean IQ between the pill group and placebo group.

   Smart pill IQ, 81, 98, 117, 76, 91, 110, 103, 105, 72, 121
   Placebo IQ, 79, 80, 89, 94, 99, 106, 111, 109, 73, 119

a) Identify the independent variable and its levels
   IV = Pill Type 
b) Is the independent variable categorical or numeric?
   IV = Nominal Categorical Variable 
c) Identify the dependent variable
   DV = IQ (Intelligence Quotient)
d) Is the dependent variable categorical or numeric?
   DV = Numeric Variable
   Note: In this case, it is a Discrete Numeric Variable. However, IQ is 
   sometimes expressed using decimals, which, in those cases, would make it a 
   Continuous Numeric Variable.

Question 2 Parts e) and f)
```{r Question 2 Parts e and f}
# e) Enter the data in R as a data frame. Submit your R Markdown file to d2l 
#    drop box

# each = 10 ensures pill type's assignments repeat ten times.
# This aligns the IQ scores with each pill type.
# rep() function replicates values in a vector.
IV_Group <- rep(c("Smart Pill", "Placebo"), each = 10)
DV_IQ <- c(81, 98, 117, 76, 91, 110, 103, 105, 72, 121, 
           79, 80, 89, 94, 99, 106, 111, 109, 73, 119)

# create data frame using the vectors
pill_data <- data.frame(IV_Group, DV_IQ)
pill_data

# f) Generate a graph to visualize the means of IQ from the smart pill group and 
#    the placebo group. Submit your R Markdown file.
ggplot(pill_data, aes(x = IV_Group, y = DV_IQ, fill = IV_Group)) +
  geom_bar(stat = "summary", fun = "mean", width = 0.3) +
  labs(title = "Comparison of Mean IQ Scores", x = "Group", y = "Mean IQ") +
  theme_minimal()
```

3. Listed below are the lead concentrations (in ug/g) measured in different 
   Ayurveda medicines. Ayurveda is a traditional medical system commonly used in 
   India. The lead concentrations listed here are from medicines manufactured in 
   the United States. The data below are based on the article “Lead, Mercury, 
   and Arsenic in US and Indian Manufactured Ayurvedic Medicines Sold via the 
   Internet,” by Saper, et al., Journal of the American Medical Association.
   
   3.0 6.5 6.0 5.5 20.5 7.5 12.0 20.5 11.5 17.5

   What is the mean, standard deviation, range and median of this sample data?
   
Question 3
```{r Question 3}
numbers <- c(3.0, 6.5, 6.0, 5.5, 20.5, 7.5, 12.0, 20.5, 11.5, 17.5)
mean(numbers)
sd(numbers)
range(numbers)
median(numbers)
```

4. The data Orange in R recorded the age and circumferences of 5 different types 
   of orange trees. The variable “Tree” recorded the five types as 1, 2, 3, 4, 
   and 5. The variable “age” recorded the ages of the trees and the variable 
   “circumference” recorded the circumferences of the trees. Use R pipeline to
   
a) Keep only tree 1, 2 and 3.
b) Find the mean and standard deviation of circumferences of each type of the 
   trees for tree type 1, 2 and 3. List the means and standard deviations:
c) Generate boxplots for tree type 1, 2 and 3. Submit your R Markdown file.

Question 4 Parts a - c
```{r Question 4 Parts a - c}
# view the data
Orange

# Part a: keep 1, 2, and 3
first_three <- Orange %>%
  filter(Tree == 1 | Tree == 2 | Tree == 3) # OR (|) operator

# Part b: find mean and standard deviation
stats <- first_three %>%
  group_by(Tree) %>%
  summarize(mean = mean(circumference), sd = sd(circumference))
stats

# Part c: generate boxplots with ordered tree types
ggplot(first_three, aes(x = factor(Tree, levels = c(1, 2, 3)), # ascending order
                        y = circumference)) +
  geom_boxplot() +
  labs(title = "Boxplot of Circumferences by Tree Type", x = "Tree Type",
       y = "Circumference (cm)") +
  theme_minimal()
```

