---
title: "Homework 2"
author: "Peyton Hall"
date: "01/29/2025"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Necessary Libraries
```{r Load Necessary Libraries}
library(tidyverse)
library(ggplot2)
library(readxl)
```

Question 1
```{r Question 1}
# The following data come from a study that examines the efficacy of saliva 
# cotinine as an indicator for exposure to tobacco smoke. In one part of the 
# study, seven subjects were each required to smoke a single cigarette. Samples 
# of saliva were taken from all individuals 12 and 24 hours after smoking the 
# cigarette and cotinine levels were recorded.

# Cotinine levels (nmol/l)
# Individuals, After 12 hours, After 24 hours
# 1,           73,             24
# 2,           58,             27
# 3,           67,             49
# 4,           93,             59
# 5,           33,             0
# 6,           18,             11
# 7,           147,            43

# Perform an appropriate test to see if the mean cotinine levels after 12 hours 
# is significantly higher than the mean cotinine level after 24 hours with 0.05 
# as the significance level.

# Step 1: Formulate the null and alternative hypotheses using symbols
# Null Hypothesis (H_0): The mean cotinine levels after 12 hours are equal to 
#                        the mean cotinine levels after 24 hour. 
# μ_12 = μ_24
# Alternative Hypothesis (H_a): The mean cotinine levels after 12 hours are 
#                               higher than the mean cotinine levels after 24 
#                               hours.
# μ_12 > μ_24

# Step 2: Choose the _____ tailed ____________test
#         Choose the one-tailed paired t-test.

# Step 3: Find the test statistic and p-value from R
after_12_hours <- c(73, 58, 67, 93, 33, 18, 147)
after_24_hours <- c(24, 27, 49, 59, 0, 11, 43)
t.test(after_12_hours, after_24_hours, alternative = "greater", paired = TRUE)

# Step 4: Make a decision based to the null hypothesis; and explain the decision 
# in the context
```
t = 3.3228; p-value = 0.007974
Reject H0; there is evidence to support that the mean cotinine levels after 12 
hours are significantly higher than the mean cotinine levels after 24 hours. 
This suggests that the cotinine concentration in saliva decreases significantly 
over time after smoking a cigarette.

Question 2
```{r Question 2}
# The Bayley Scales of Infant Development yield scores on the Psychomotor 
# Development Index (PDI). PDI can be used to assess a child’s level of 
# functioning at approximately one year of age. As part of a study investigating 
# the development status of children who had undergone reparative heart surgery 
# during the first three months of life, the Bayley Scales were administered to 
# a sample of one-year-old infants born with congenital heart disease. The
# children had been randomized to one of two different treatment groups, known 
# as “circulatory arrest” (denoted as treatment group 0) and “low-flow bypass” 
# (denoted as treatment group 1). The groups differed in the specific way in 
# which the reparative surgery was performed. Unlike circulatory arrest, 
# low-flow bypass maintains continuous circulation through the brain; although 
# it is felt to be preferable by some physicians, it also has its own associated 
# risk of brain injury. The data is on d2l (PDI score). Assume the two 
# treatments have unequal variances. At the 0.05 level of significance, test the 
# claim that the mean PDI score at one year of age for the circulatory arrest 
# treatment group is significantly different from that for the low-flow group.

PDI_data <- read_excel("~/Desktop/STAT 301/Week 2/PDI data.xlsx")
PDI_data

# Step 1: Formulate the null and alternative hypotheses using symbols
# Null Hypothesis (H_0): The mean PDI score at one year of age for the 
#                        circulatory arrest treatment group (TG0) is equal to 
#                        that for the low-flow group (TG1).
# μ_TG0 = μ_TG1
# Alternative Hypothesis (H_a): The mean PDI score at one year of age for the 
#                               circulatory arrest treatment group (TG0) is not 
#                               equal to that for the low-flow group (TG1).
# μ_TG0 ≠ μ_TG1

# Step 2: Choose the ______tailed ___________test
#         Choose the two-tailed t-test.

# Step 3: Find the test statistic and p-value from R

pdi_tg0 <- PDI_data$PDI[PDI_data$Treatment == 0]  # Data for circulatory arrest
pdi_tg1 <- PDI_data$PDI[PDI_data$Treatment == 1]  # Data for low-flow bypass
# perform the Welch t-test for unequal variances
t.test(pdi_tg0, pdi_tg1, alternative = "two.sided", var.equal = FALSE)

# Step 4: Make a decision to H0, and explain the decision in terms of the 
# original claim.

# Which group has a higher PDI score? Low-flow bypass group (TG1).
```
t = -2.1715; p-value = 0.03158
Reject H0; there is evidence to support that the mean PDI score at one year of 
age for the circulatory arrest treatment group is significantly different from 
that for the low-flow bypass group.

Question 3
```{r Question 3}
# A camera has been developed to detect the presence of cataract more 
# accurately. Using the camera, the gray level of each point in the lens of a 
# human eye can be characterized into 256 gradations. To test the camera, 
# photographs were taken of six randomly selected normal eyes and six randomly 
# selected cataractous eyes from a different group of people. The gray level of
# each eye was computed in the lens. The data are given in the following table. 
# Use the 0.05 significance level to test the claim that the mean gray level 
# from cataractous eyes is significantly lower than the mean gray level from 
# normal eyes.

# Cataractous gray level, Normal gray level
# 161,                    158
# 140,                    182
# 136,                    185
# 161,                    145
# 106,                    167
# 149,                    177

# a) Generate a point graph to plot the mean difference between cataractous eyes 
# and normal eyes with error bars. Which one has a higher mean gray level?
eye_data <- data.frame(
  Cataractous = c(161, 140, 136, 161, 106, 149),
  Normal = c(158, 182, 185, 145, 167, 177)
)

eye_data$Difference = eye_data$Normal - eye_data$Cataractous
summary_stats <- eye_data %>%
  summarise(Mean_Difference = mean(Difference), SD = sd(Difference))
summary_stats

means <- colMeans(eye_data[, c("Cataractous", "Normal")])
errors <- apply(eye_data[, c("Cataractous", "Normal")], 2, sd) / sqrt(nrow(eye_data))

error_bars <- data.frame(
  Group = c("Cataractous", "Normal"),
  Mean = as.numeric(means),
  Error = as.numeric(errors)
)

ggplot(error_bars, aes(x = Group, y = Mean, ymin = Mean - Error, ymax = Mean + Error)) +
  geom_point() +
  geom_errorbar(width = 0.2) +
  labs(title = "Mean Gray Level Comparison with Error Bars",
       x = "Eye Type",
       y = "Mean Gray Level") +
  theme_minimal()

# b) Test the hypothesis in the question with the steps below:

# Step 1: Formulate the null and alternative hypotheses using symbols
# Null Hypothesis (H0): The mean gray level from cataractous eyes is equal to 
#                       the mean gray level from normal eyes.
# μ_cataractous = μ_normal
# Alternative Hypothesis (H_a): The mean gray level from cataractous eyes is 
#                               less than the mean gray level from normal eyes.
# μ_cataractous < μ_normal

# Step 2: Choose the ______tailed ___________test
#         Choose the one-tailed t-test

# Step 3: Find the test statistic and p-value from R
t.test(eye_data$Cataractous, eye_data$Normal, alternative = "less")

# Step 4: Make a decision to H0, and explain the decision to address the 
# original cl
```
t = -2.5597; p-value = 0.015
Reject H0; there is evidence to support that the mean gray level from 
cataractous eyes is significantly lower than the mean gray level from normal 
eyes.

Question 4
```{r Question 4}
# Data below lists the times required for randomly selected eight flights to 
# taxi out for takeoff and the corresponding times required to taxi in after 
# landing for the same flight. All eight flights are American Airlines from New 
# York to Los Angeles. Use the significance level of 0.01 to test the claim that 
# the average time to taxi in is different from the average time to taxi out.

# Flight,    1,  2,  3,  4,  5,  6,  7,  8
# Taxi-out, 30, 19, 12, 19, 18, 22, 19, 13
# Taxi-in,  12, 13,  8, 21, 17, 15, 26, 12

# Step 1: State the null and alternative hypotheses using the correct symbols
# Null Hypothesis (H_0): The mean taxi-in time (μ_in) is equal to the mean 
#                        taxi-out time (μ_out).
# μ_in = μ_out
# Alternative Hypothesis (H_a): The mean taxi-in time (μ_in) is not equal to the 
#                               mean taxi-out time (μ_out).
# μ_in ≠ μ_out

# Step 2: Choose the ____tailed ________________ test
#         Choose the two-tailed paired t-test.

# Step 3: Find the test statistic and p-value in R
taxi_out <- c(30, 19, 12, 19, 18, 22, 19, 13)
taxi_in <- c(12, 13, 8, 21, 17, 15, 26, 12)
t.test(taxi_in, taxi_out, paired = TRUE, alternative = "two.sided")
mean(taxi_out)
mean(taxi_in)

# Step 4: Make a decision to the null hypothesis; and explain the decision to 
# address the original claim.
```
t = -1.3401; p-value = 0.2221
Fail to reject H0; there is no evidence to support that the average time to taxi 
in is different from the average time to taxi out.
Based on their means (taxi out being 19 and taxi in being 15.5), there is a 
difference, yet that difference is not statistically significant due to the 
significance level (α = 0.01).
