---
title: "Worksheet2"
author: "Peyton Hall"
date: "01/24/2025"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Necessary Libraries
```{r Load Necessary Libraries}
library(readxl)
library(tidyverse)
```

$H_0: μ = 80$ vs $H_a: μ < 80$
In Class Coding One Sample T Test
```{r In Class Coding One Sample T Test}
Reading <- read_excel("~/Desktop/STAT 301/Week 2/Reading.xlsx")
t.test(Reading$read, mu = 80, alternative = "less")
```
t = -3.87, p-value = 0.0003 
Reject H0; There is evidence to support that the mean reading score from Prof.
White’s students is significantly lower than 80

$H_0: μ = 0.5$ vs $H_a: μ > 0.5$
Worksheet 2 Question 2
```{r Worksheet 2 Question 2}
Mehr_Song_and_Spelke <- read_excel("~/Desktop/STAT 301/Week 2/Mehr Song and Spelke.xlsx")

M1 <- Mehr_Song_and_Spelke %>%
  filter(exp1 == 1)

t.test(M1$Baseline_Proportion_Gaze_to_Singer, mu = 0.5, alternative = "greater")
```
t = 0.67; p-value = 0.25
Fail to reject H0; There is no evidence to support that the average proportion 
of infants looking for the singer was higher than 0.5.

$H_0: μ(drug) = μ(placebo)$ vs $H_a: μ(drug) > μ(placebo)$
In Class Coding Slide 11
```{r In Class Coding Slide 11}
drugtype <- rep(c("drug", "placebo"), c(10,12))

birthwt <- c(6.9, 7.6, 7.3, 7.6, 6.8, 7.2, 8.0, 5.5, 5.8, 7.3,
             6.4, 6.7, 5.4, 7, 5.3, 6.6, 5.8, 5.7, 6.2, 7.1, 5.6, 4.2)
babydf <- data.frame(drugtype, birthwt)
babydf
```

```{r}
library(ggplot2)
newdf <- babydf %>%
  group_by(drugtype) %>%
  summarize(ave = mean(birthwt), ste = sd(birthwt)/sqrt(length(birthwt)),
            tstar = qt(1-0.05/2, length(birthwt) - 1))
ggplot(data = newdf, aes(x = drugtype, y = ave) )+ geom_point() + geom_errorbar(aes(ymin = ave - tstar * ste, ymax = ave + tstar * ste)) + ylim(c(0,8))
```

$H_0: μ(drug) = μ(placebo)$ vs $H_a: μ(drug) > μ(placebo)$
In Class Coding Slide 11 T Test
```{r In Class Coding Slide 11 T Test}
# syntax implies μ(drug) > μ(placebo)
t.test(birthwt~drugtype, data = babydf, alternative = "greater")
```
t = 2.87; p-value = 0.0047
Reject H0; there is evidence to support that the average baby weight from the 
drug group is significantly higher than the average baby weight from the placebo 
group

```{r}
airline <- rep(c("delta", "sun"), each = 9)

fuel <- c(82.1, 68.4, 52.7, 58.8, 52, 71.4, 60, 59.6, 44.9,
          60.8, 68.3, 77.2, 76.4, 71.4, 63, 81.1, 72.1, 73.3)
airdf <- data.frame(airline, fuel) 
```

```{r}
airdf %>%
  group_by(airline) %>%
  summarize(ave = mean(fuel), ste = sd(fuel)/sqrt(length(fuel)), tstar = qt(1-0.05/2, length(fuel)-1)) %>%
  ggplot(aes(x=airline, y= ave)) + geom_point() + geom_errorbar(aes(ymin=ave - tstar*ste, ymax=ave+tstar*ste))
```

```{r}
t.test(fuel~airline, data=airdf, alternative = "two.sided")
```
Reject H_0
There is evidence to support that the ave. fuel efficiency differs sig. between 
2 airlines